A neural coarse graining theory of consciousness

### 2018-07-26 Paper_neural_coarse_graining_of_consciousness.tex
ATOM try

@@

			%% scale

@@

			%% our argument
				@ In this article, we propose that consciousness \highlightAndComment[YellowGreen]{correlates}{Is this proper?} information processing at a certain coarse-grained level where the neural system creates a high degree of non-trivial informational closure. We further predict that \textit{levels} of consciousness are determined by the levels of informational closure and \textit{contents} of consciousness are determined by the concurrent states of the informational closure \citep{laureys2005neural, overgaard2010neural}.\todo{There must be a better way to reword this part}

			%% About this paper
				@ In the following sections, we first introduce non-trivial informational closure and its importance to information processing for human scale agents.

				@ We argue that through coarse-graining the human neural system can achieves non-trivial informational closure at a specific coarse-grained level.

				@ We then propose that the state of the informational closure correlates the state of consciousness for both levels and contents which can be explained and computationally quantified and predicted in this theory.

				@ We discuss how this theory can concisely explain empirical results from previous studies and reconcile several current major theories of consciousness.

				@ We finally provide predictions and testable approaches for this theory.

@@

%		==================================================
%		\begin{ants}
%			
%			@ consciousness is information
%			
%			@ conscious contents
%				@@ \toWrite{lower level}
%				@@ Information processing of individual neurons are noisy.
%				@@ We don't have conscious content changes when individual neurons fire action potentials. 				
%				@@ Don't even mention ion channel dynamic. 
%				
%				
%				@@ \toWrite{Higher-level dynamic}
%
%				
%				
%				
%				
%			
%			@ For human, physics at our body scale are nearly follow classic physics, which is deterministic. 
%			
%			
%			@ \toWrite{Human tendency for active prediction }
%				@@ \cite{Stahl2015} \rewrite{
%					Given the overwhelming quantity of information available from the environment, how do young learners know what to learn about and what to ignore? We found that 11-month-old infants (N = 110) used violations of prior expectations as special opportunities for learning. The infants were shown events that violated expectations about object behavior or events that were nearly identical but did not violate expectations. The sight of an object that violated expectations enhanced learning and promoted information-seeking behaviors; specifically, infants learned more effectively about objects that committed violations, explored those objects more, and engaged in hypothesis-testing behaviors that reflected the particular kind of violation seen. Thus, early in life, expectancy violations offer a wedge into the problem of what to learn.}
%			
%			@ Intuitive physics \needref{Intuitive physics}
%			
%			@ However, sensory signal are always noisy. 			
%			
%			
%			@ We propose that consciousness is not related to particular neural architectures or location of the neural system.
%			
%
%		\end{ants}

	\section{Non-trivial informational closure\martin{todo}}

		\subsection{Meaning of informational closure}
			\begin{ants}
				@ \rewrite{
					The notion of closure plays a prominent role in systems theory where it is used to identify or define the system in distinction from its environment} \cite{BERTSCHINGER.2006}

@@

			\noindent However, achieving information closure (minimising $J_t$) could be trivial, which means the environment and the system are entirely independent of each other.
				\begin{equation}
				\begin{aligned}
				{I(Y_{t+1};E_{t})=0}&&{\Rightarrow}&&{J_{t}(E \rightarrow Y )=0}
				\end{aligned}
				\end{equation}

			\noindent In the non-trivial case, the system trying to encode environmental dynamic can also achieve informational closure. This means
				\begin{equation}
				I(Y_{t+1};E_{t})\neq0, 
				\end{equation}

			\noindent And, non-trivial informational closure can be defined as 		
				\begin{equation}
				\left.\begin{array}
				{rl}{NTIC} & {:=I(Y_{t+1};Y_{t})-I(Y_{t+1};Y_{t}|E_{t})}\\
				{ } & {\ =I(Y_{t+1};E_{t})-I(Y_{t+1};E_{t}|Y_{t})}
				\end{array} \right.
				\end{equation}

			\noindent Therefore, Maximising NTIC is equivalent to
				\begin{equation}
				\begin{aligned}
				& \text{maximise} & { } & I(Y_{t+1};Y_{t}) & { } & \text{and} \\
				& \text{minimise} & { } & I(Y_{t+1};Y_{t}|E_{t}) & { }
				\end{aligned}
				\end{equation}

		\subsection{Trivial case}
			\begin{ants}
				@ However, informational closure could be trivial.

				@ Information closure could be trivial. That is, when a system is fully independent from the environment, no information can flow into the system from the environment and also leak to the environment from the system. Equa. xxxx shows the mathematical description of trivial informational closure. When mutual information between the environment and system future state close to 0, and the system transition is independent from the environment, the system can reach informational closure. However, such systems do not have any functional meaning and evolutional advantages.

				@ \rewrite{
					A system that is independent from its environment trivially achieves informational closure.}  \cite{BERTSCHINGER.2006}
					\begin{equation}
						\left.\begin{array}
						{l}{MI(S_{n+1};E_{n})=0}\\
						{MI(S_{n+1};S_{n})-MI(S_{n+1};S_{n}|E_{n})=0}
						\end{array}\right.
					\end{equation}
			\end{ants}

		\subsection{Non-trivial informational closure}
			\begin{ants}

				@ \citep{BERTSCHINGER.2006}
				@ \citep{guttenberg2016neural}

				@ \rewrite{The “informational closure” becomes non-trivial if the state contains information about the environment, i.e. $MI(S_{n+1};E_{n})\neq0$}

				@ We can define non-trivial informational closure (NTIC)

				@ Definition:\callforhelp{Which def is better? }

					@@
					\begin{equation}
					N T I C = H ( Y _ { t + 1} ) - H ( Y _ { t + 1} | Y _ { t } )
					\end{equation}
					\citep{guttenberg2016neural}

					@@
					\begin{equation}
					\left.\begin{array}{l}{ N T I C  : = M I ( S _ { n + 1} ; E _ { n } ,\dots ,E _ { n - m } ) - M I ( S _ { n + 1} ; E _ { n } ,\dots ,E _ { n - m } | S _ { n } ) }\\{  = M I ( S _ { n + 1} ; E _ { n } ,\dots ,E _ { n - m } ) - M I ( S _ { n + 1} ; E _ { n } | S _ { n } ) }\end{array} \right.
					\end{equation}
					\cite{BERTSCHINGER.2006}

				@ \rewrite{Modeling: The system reaches synchronization and internalizes the correlations observed in the environment by building up own structures.}

		\subsection{Achieve non-trivial informational closure}
			\begin{ants}

				@ Accurately model deterministic environment 
				@ \rewrite{
					This demonstrates that a system exhibiting certain internal regularities as measured by $A^* = MI(Sn+1; Sn)$ can achieve informational closure either by gaining information about the environment or by increased autonomy, i.e. by becoming unpredictable or uncontrollable from the (13) environment. Therefore, information about the environment, i.e. modeling, and autonomy can be considered as complementary strategies for achieving informational closure.}

@@

				@ In the next section, we claim that the neural system achieve NTIC through coarse-graining.

@@

	\section{Neural coarse-graining}

			%% About coarse-graining
%			@ \rewrite{
%				Coarse-graining: a method of reducing the complexity of a system by treating groups of atoms/molecules as single quasi-particles CG: coarse-grained}
%			
%			
%			@ \rewrite{
%				a number of different microstates,First, typically a number of different microstates, will realize the same value of the macro-variables of upper level causal theories.}
%			
%			
%			@ One consequence of coarse-graining is that it makes it permissible to ignore certain causal factors that would be relevant at more fine-grained level of description. \cite{price2007causation}

@@

		\subsection{Coarse-graining in neural system}
			\begin{ants}

@@

@@

					@@ firing rate coding

@@

@@

	\section{A neural coarse graining theory of consciousness}
		\ideaBox{Need to describe my intuition more concretely}

@@

			@

@@

			%% reasoning
				@ Information processing in the neural system exists at different level of abstraction.

@@

			%% our claim

@@

>>>>

% The name of our theory
\newcommand{\tnamefull}{[TheoryNameFull]}
\newcommand{\tname}{[TheoryNameShort]}
\newacro{\tname}{\tnamefull}

@@

			%% scale
				@ ============ About scale ===========

@@

			%% our argument
				@ =========== our argument ===========

				@ In this article, we propose new theory of consciousness, called \ac{\tname}, which states that consciousness \highlightAndComment[YellowGreen]{correlates}{Is this proper?} information processing at a certain coarse-grained level where the neural system creates a high degree of non-trivial informational closure (NTIC). We further postulate that the \textit{level} of consciousness corresponds to the degree of NTIC and the \textit{contents} of consciousness corresponds to the state of the NTIC process (for level versus content of consciousness see \cite{laureys2005neural, overgaard2010neural}).

			%% About this paper
				@ ========= About this paper =========

				@ Since this theory rests on two core concepts: non-trivial informational closure and neural coarse-graining, in the following sections, we first introduce non-trivial informational closure and its importance to information processing for human scale agents.

				@ We next argue that through coarse-graining the human neural system can achieves non-trivial informational closure at a specific coarse-grained level.

				@ We then propose that the state of the informational closure correlates the state of consciousness for both levels and contents which can be explained and computationally quantified and predicted by this theory.

				@ We also address how this theory can concisely explain empirical results from previous studies and reconcile several current major theories of consciousness.

				@ Finally, we provide predictions and testable approaches for this theory.

@@

		%% backup point
		\begin{comment}			
			@ consciousness is information

			@ conscious contents
				@@ \toWrite{lower level}
				@@ Information processing of individual neurons are noisy.
				@@ We don't have conscious content changes when individual neurons fire action potentials. 				
				@@ Don't even mention ion channel dynamic.

				@@ \toWrite{Higher-level dynamic}

			@ For human, physics at our body scale are nearly follow classic physics, which is deterministic.

			@ \toWrite{Human tendency for active prediction }
				@@ \cite{Stahl2015} \rewrite{
					Given the overwhelming quantity of information available from the environment, how do young learners know what to learn about and what to ignore? We found that 11-month-old infants (N = 110) used violations of prior expectations as special opportunities for learning. The infants were shown events that violated expectations about object behavior or events that were nearly identical but did not violate expectations. The sight of an object that violated expectations enhanced learning and promoted information-seeking behaviors; specifically, infants learned more effectively about objects that committed violations, explored those objects more, and engaged in hypothesis-testing behaviors that reflected the particular kind of violation seen. Thus, early in life, expectancy violations offer a wedge into the problem of what to learn.}

			@ Intuitive physics \needref{Intuitive physics}

			@ However, sensory signal are always noisy.

			@ We propose that consciousness is not related to particular neural architectures or location of the neural system.			
		\end{comment}

	\section{Non-trivial informational closure\martin{todo}}
		\ideaBox{Not easy. Should we go for NTIC without describe an intuition about consciousness? Or we should start from intuition? Or explain the intuition in Chapter 4?}

		\subsection{Meaning of informational closure}
			\begin{ants}
				@ In systems theory the concept of closure \citep{maturana1991autopoiesis, rosen1991life, pattee2012evolving, luhmann1995probleme}

				@ The notion of non-trivial informational closure (NTIC) is proposed by \cite{BERTSCHINGER.2006}.

				@ The concept of closure is closely related to system identification in systems theory. One can define a system from its environment by computing the closureness of the system.

				a prominent role in systems theory where it is used to identify or define the system in distinction from its environment

				@ Consciousness as an intrinsic information. Therefore, information should be self-define. This means the boundary of conscious content should be related to informational closure.

				@ ============================================================================

				@ \rewrite{
					The notion of closure plays a prominent role in systems theory where it is used to identify or define the system in distinction from its environment} \cite{BERTSCHINGER.2006}

@@

		% trivial
			\noindent 
			Achieving information closure (minimising $J_t$) could be trivial, which means the environment and the system are entirely independent of each other.
				\begin{equation}
				\begin{aligned}
				{I(Y_{t+1};E_{t})=0}&&{\Rightarrow}&&{J_{t}(E \rightarrow Y )=0}
				\end{aligned}
				\end{equation}

		% non-trivial
			\noindent 
			However, informational closure can be achieved non-trivially. In the non-trivial case, the system trying to encode information about the environmental dynamics can also be informational closured. This means
				\begin{equation}
				I(Y_{t+1};E_{t}) > 0
				\end{equation}

			\noindent
			This also implies
				\begin{equation}
					I(Y_{t+1};Y_{t})-I(Y_{t+1};Y_{t}|E_{t}) > 0
				\end{equation}

			\noindent Therefore, non-trivial informational closure can be defined as 		
				\begin{equation}
				\left.\begin{array}
				{rl}{NTIC} & {:=I(Y_{t+1};Y_{t})-I(Y_{t+1};Y_{t}|E_{t})}\\
				{ } & {\ =I(Y_{t+1};E_{t})-I(Y_{t+1};E_{t}|Y_{t})}
				\end{array} \right.
				\end{equation}

			\noindent And, Maximising NTIC is equivalent to
				\begin{equation}\label{eq:nticObjective}
				\begin{aligned}
				& \text{maximise} & { } & I(Y_{t+1};Y_{t}) & { } & \text{and} \\
				& \text{minimise} & { } & I(Y_{t+1};Y_{t}|E_{t}) & { }
				\end{aligned}
				\end{equation}

			\toWrite{need to elaborate this part more}
			When a system achieve NTIC, the system internalizes and synchronises the dynamics of the environment, which also means modelling the environment. Having high degrees of NTIC entails predictive power about the environment. This gives biological agents a great functional meaning and an evolutional reason reason to achieve NTIC in their information processing systems. Practically, \cite{guttenberg2016neural} proposed that NTIC can be achieved by maximising the predictive power about the environment and the self-predictive information of the system concurrently . \callforhelp{Do we need to say we also maximise entropy of latent layers? }

		\subsection{Biological evidence of informational closure in the neural system}
			\begin{figure}[H]
				\includegraphicsTodo[width=\textwidth]{WritingMaterials/Palmer2015.png} 
				\caption{Figure adapted from \cite{sederberg2018learning}}
				\label{fig:sederberg2018learning}	
			\end{figure}

			\begin{ants}
				@ So far, not much studies directly examine NTIC in biological systems.

				@ Some very recent studies have showed relevant properties.\todo{Perhaps we should put a picture about this study here.}

				@ By examining salamander retina, \cite{Palmer2015, sederberg2018learning} found that neural populations of retinal ganglion cells encoded predictive information about external stimuli also had high self-predictive information about their own future states on two different classes of stimuli.

				@ This property matches Eq. (\ref{eq:nticObjective})

				@ Therefore, self-predictive information of the neural populations can guides  prediction about external stimuli without reference to stimulus parameters explicitly.

				@ efficient readouts of one type of information were likely to be efficient readouts of the other. The generalization of internal predictability across stimulus classes partially reflects the ability of the retina to adapt to the statistics of stimuli.

				@ \rewrite{ Internal predictability reflects the timescale of correlation in the stimulus }

				@ The two studies provides a empirical evidence that NTIC may serve as an objective in the neural system. 
			\end{ants}

			\begin{writingMaterials}
				@ ===

				@ \rewrite{
					Internal information reflects the spatial and temporal correlations in the stimulus.}

				@ \rewrite{predictive information is encoded synergistically.}

				@ \rewrite{
					in recordings of populations of RGCs of salamander retina driven by a simple stimulus with partially predictable dynamics, joint activity patterns transmit information that is predictive of future stimuli }

			%% backup
			\begin{comment}
				@ Accurately model deterministic environment 
				@ \rewrite{
					This demonstrates that a system exhibiting certain internal regularities as measured by $A^* = MI(Sn+1; Sn)$ can achieve informational closure either by gaining information about the environment or by increased autonomy, i.e. by becoming unpredictable or uncontrollable from the (13) environment. Therefore, information about the environment, i.e. modeling, and autonomy can be considered as complementary strategies for achieving informational closure.}

@@

				@ In the next section, we claim that the neural system achieve NTIC through coarse-graining.				
			\end{comment}

@@

	\section{Neural coarse-graining}
		\subsection{intro}
			\begin{ants}
				@ As we mentioned above, to construct the NTIC process can bring agents great survival advantage.

				@ However, to achieve NTIC (or even IC) is challenging for highly stochastic systems. Due to the stochasticity, the current states of the system have low predictability of its future states.

				@ Previous research has been showing that the neural system are suffered from multiple noise sources  during information processing, including sensory, cellular, electrical, synaptic noises and have large trial-to-trial variability at the cellular level. \citep{faisal2008noise}. neurons are noisy: the same pattern of activity never occurs twice, even when the same stimulus is presented.

				@ These noises sabotaged the predictability of future states and prevent the neural system from creating NTIC processes at the cellular level.

				@ And every individual sensor only receives a little amount of information about the environment.

				@ We argues that the human neural system to achieve NTIC through coarse-graining. 
			\end{ants}

			%% About coarse-graining			
			@ One consequence of coarse-graining is that it makes it permissible to ignore certain causal factors that would be relevant at more fine-grained level of description. \cite{price2007causation}

@@

		\subsection{Coarse-graining in neural system}
			\begin{ants}
				@

				@ \rewrite{Neurons are subject to thermal fluctuations and other physical noise (Faisal et al., 2008).} \cite{faisal2008noise}

@@

				@ \rewrite{Because the sequence of action potentials generated by a given stimulus varies from trial to trial, neuronal responses are typically treated statistically or probabilistically}

@@

					@@ Rate coding
						\cite{stein2005neuronal}

@@

					@@ \rewrite{
						Averaging is used in many neural systems in which information is encoded as patterns of activity across a population of neurons that all subserve a similar function (for example, see REFs 140,141 )}
										140. Georgopoulos, A., Schwartz, A. \& Kettner, R. Neuronal population coding of movement direction. Science 233, 1416–1419 (1986). 141. Lee, C., Rohrer, W. H. \& Sparks, D. L. Population coding of saccadic eye movements by neurons in the superior colliculus. Nature 332, 357–360 (1988).

					@@ \rewrite{
						Electrophysiological studies in the monkey have shown that behaviourally relevant signals are averaged not only across neuronal populations but also over time in the formation of a behavioural decision 156 .} 156. Gold, J. I. \& Shadlen, M. N. Representation of a perceptual decision in developing oculomotor commands. Nature 404, 390–394 (2000).

					@@ \rewrite{A distributed representation of information of this type is more robust to the effects of noise.}

@@

		\subsection{NTIC at specific levels of neural coarse-graining\callforhelp{any reference?}}
			\ideaBox{Do we need to discuss Hoel's theory here?}
			\begin{ants}
				@ The relation between the degrees of NTIC and coarse-graining is not monotonic.

				@ Overly coarse-graining a system can lead to low degrees of NTIC.

				@ Imagining an extreme scenario that all microscopic states are mapped to a macroscopic state, the macroscopic process losses all the information can be used to make self-prediction.

				@ \cite{sederberg2018learning} showed similar results in salamander retina neural populations.

				@ They found that

				@ \rewrite{
					We found that while the efficiency of the learned readouts relative to optimal single-bit readouts remained high the efficiency relative to the full input group decreased, such that readouts of groups of 10 cells typically preserved 30\% of the total predictive information}

				@ \rewrite{
					In other words, reading out from groups of four cells can be accomplished with high absolute efficiency. Perhaps the best way to combine more than four cells is to break down the readout into indivisible units of four cells, which are later recombined in subsequent processing.}
			\end{ants}

	% ============================================================================ %
	%                                  Out theory                                  %
	% ============================================================================ %
	\section{A neural coarse graining theory of consciousness}
		\ideaBox{Need to describe my intuition more concretely}

@@

			@ Therefore, to understand this system, you don't need more information from external environment.

			@ It also means everything in this system is self-defined.

@@

			%% reasoning

				@ ============= reasoning ============

				@ Information processing in the neural system exists at different level of abstraction.

@@

			%% our claim 
				@ ============ our claim  ============

@@

		\subparagraph{more information theory and philosophy relevance}
			\begin{writingMaterials}
				@ One consequence of coarse-graining is that it makes it permissible to ignore certain causal factors that would be relevant at more fine-grained level of description. \cite{price2007causation}

				@ This means the fine-grain variable does not exist at the coarse-grained causal chain. 
			\end{writingMaterials}
